---
title: "Exercise set 1"
author: "Joseba Hernandez Bravo and Jorge Vicente Puig"
date: "17/12/2021"
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

```{r, message = FALSE}
library(combinat)
```

In this first task we will examine the correlation between the pairs of (X1,Y1)
values by means of a correlation test. 

Firstly we will define two different null hypothesis.

- $H_{0}:\rho=0$: There is no correlation between Chest circumference and volume of air
- $H_{0}:\rho>0$: Chest circumference and volume of air are positively correlated

# $H_{0}:\rho = 0$ 

```{r}
x= c(39,29,60,40,32)
y=c(11,5,20,8,6)

sttrue= cor(x,y)

n=length(y)
nr=fact(n) #number of rearrangements to be examined
st=numeric(nr)

cnt=0
d=permn(y)

for (i in 1:nr){
  st[i]<-cor(d[[i]],x)
  if (st[i] >=sttrue)
    cnt=cnt+1 
}
```

```{r}
print(paste("p-value= ",cnt/nr))
print(paste("Null hypothesis: ", sttrue))
```
```{r}
hist(st)
abline(v=sttrue,col="blue",lwd=2)
```
As the p values is less than the significance lece ($0.05$) we will reject the null
hypothesis. So, we can say that there is sufficient evidence to conclude that the 
relationship between X an Y could be linear. 
# $H_{0}:\rho > 0$ 



